### 2018-05-17 AppendixI.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Article EcoFoG
% Version 2.1 (23/10/2017)
%
% adapté de :
% Stylish Article
% LaTeX Template
% Version 1.0 (31/1/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Mathias Legrand (legrand.mathias@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,10pt]{ArtEcoFoG} % Document font size and equations flushed left

\setcounter{tocdepth}{3} % Show only three levels in the table of contents section: sections, subsections and subsubsections

% Pandoc environments
\usepackage{framed}
\usepackage{fancyvrb}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, fontsize=\scriptsize} % Code R
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{caption}
% These lines are needed to make table captions work with longtable:
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
% longtable 2 columns
% https://tex.stackexchange.com/questions/161431/how-to-solve-longtable-is-not-in-1-column-mode-error
\makeatletter
\let\oldlt\longtable
\let\endoldlt\endlongtable
\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
\def\longtable@i[#1]{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt[#1]
}
\def\longtable@ii{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt
}
\def\endlongtable{\endoldlt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\usepackage{tabu}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\JournalInfo{Hal 00679993} % Journal information
\Archive{\ }

\PaperTitle{Appendix I} % Article title

\Authors{
Ariane MIRABEL\textsuperscript{1*}\\ Eric MARCON\textsuperscript{1}\\ Bruno HERAULT\textsuperscript{2}
} % Authors
\affiliation{
\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Ploytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
}
\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author

\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{

}

%----------------------------------------------------------------------------------------

\begin{document}

\selectlanguage{english}

\flushbottom % Makes all text pages the same height

\maketitle % Print the title and abstract box

\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Appendix figures}\label{appendix-figures}

The following information are available for this article:

\textbf{Figure A1}: Whenever the family or genus were recorded for a
vernacular name, the pool of species sampled to replace the vernacular
name was restricted to those pertaining to the same higher taxonomic
level (i.e species of the same genus when the species is missing,
species of the same family when the genus is missing)
\ref{fig:FigBayesianScheme}.

\textbf{Figure A2}: Scheme of the redundancy measurement process

\textbf{Figure A3}: Trajectories of seed mass classes proportions over
30 years after disturbance, corresponding to the median (solid line) and
0.025 and 0.975 percentile (gray envelope) observed after 50 iteration
of the taxonomic uncertainty propagation. No gap filling process was
applied in this case. Initial treatments are represented by solid lines
colors with green for control, blue for T1,orange for T2 and red for T3
\ref{fig:FigSeedMassClasses}.

\textbf{Figure A4}: Trajectories of the difference to the 1989
inventories (5 years after disturbance) over 30 years after disturbance
of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and
\textbf{(c)} Simpson diversities. Trajectories correspond to the median
(solid line) and 0.025 and 0.975 percentile (gray envelope) observed
after 50 iteration of the taxonomic uncertainty propagation. Initial
treatments are represented by solid lines colors with green for control,
blue for T1,orange for T2 and red for T3 \ref{fig:DivTaxo}.

\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation. Initial treatments
are represented by solid lines colorswith green for control, blue for
T1,orange for T2 and red for T3.and the missing trait value filling
processes. \ref{fig:DivFun}.

\textbf{Figure A6}: Summarizing plots of the functional PCA analysis
defining communities functional space. \textbf{(a)} Plot of the samples
from the BRIDGE functional database in the two dimensional main plan.
\textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)}
Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem
traits, and maximum species height). @ref(fig:PCA\_graphs).

\textbf{Figure A7}: Trajectories of the functional redundancy over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation and the missing trait
value filling processes. Initial treatments are represented by solid
lines colors with green for control, blue for T1,orange for T2 and red
for T3. \ref{fig:RedFun}.

\newpage

\section{\texorpdfstring{\textbf{Figure A1}: Bayesian process
sheme}{Figure A1: Bayesian process sheme}}\label{figure-a1-bayesian-process-sheme}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{ExternalFig/BayesianSheme}

}

\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A2}: Redundancy
metric}{Figure A2: Redundancy metric}}\label{figure-a2-redundancy-metric}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{ExternalFig/RedundancyMetric}

}

\caption{Scheme of the redundancy measurement process}\label{fig:RedMM}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A3}: Proportion of seed mass
classes along
time}{Figure A3: Proportion of seed mass classes along time}}\label{figure-a3-proportion-of-seed-mass-classes-along-time}

\begin{figure*}

{\centering \includegraphics[width=0.8\linewidth]{AppendixI_files/figure-latex/FigSeedMassClasses-1}

}

\caption{Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. No gap filling process was applied in this case. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:FigSeedMassClasses}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A4}: Richness, Shannon and
Simpson taxonomic diversity
trajectories}{Figure A4: Richness, Shannon and Simpson taxonomic diversity trajectories}}\label{figure-a4-richness-shannon-and-simpson-taxonomic-diversity-trajectories}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/DivTaxo-1}

}

\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A5}: Rao diversity trajectories
with all plots
display}{Figure A5: Rao diversity trajectories with all plots display}}\label{figure-a5-rao-diversity-trajectories-with-all-plots-display}

\begin{figure*}

{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFun-1}

}

\caption{Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A6}: Functional
PCA}{Figure A6: Functional PCA}}\label{figure-a6-functional-pca}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/PCA_graphs-1}

}

\caption{Summarizing plots of the functional PCA analysis defining communities functional space. \textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan. \textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A7}: Overall functional
redundancy}{Figure A7: Overall functional redundancy}}\label{figure-a7-overall-functional-redundancy}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/RedFun-1}

}

\caption{Trajectories of the functional redundancy over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:RedFun}
\end{figure*}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension the must be eliminated
\filename@parse{}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother

%----------------------------------------------------------------------------------------

\end{document}